# Data structures and functions {#datastructuresfunctions}

## Data structures
Package **meteoland** assists in the estimation of the following variables over lanscapes (units in parentheses):

+ `DOY`: Day of the year ([1-365]).
+ `MeanTemperature`: Mean daily temperature (in degrees Celsius).
+ `MinTemperature`: Minimum daily temperature (in degrees Celsius).
+ `MaxTemperature`: Maximum daily temperature (in degrees Celsius).
+ `Precipitation`: Daily precipitation (in mm of water).
+ `MeanRelativeHumidity`: Mean daily relative humidity (in percent).
+ `MinRelativeHumidity`: Minimum daily relative humidity (in percent).
+ `MaxRelativeHumidity`: Maximum daily relative humidity (in percent).
+ `Radiation`: Incoming radiation (in MJ/m2).
+ `WindSpeed`: Wind speed (in m/s).
+ `WindDirection`: Wind direction (in degrees from North).
+ `PET`: Potential evapo-transpiration (in mm of water).

The package deals with three kinds of spatial structures: individual points, a set of pixels from a spatial grid and full (i.e. complete) grids. The package includes six S4 spatial classes, which are defined as children of classes in package **sp**. 

### Topography
Three classes are defined to represent the variation of topographic features (i.e., elevation, slope and aspect) over space:

+ Class `SpatialPointsTopography` extends `SpatialPointsDataFrame` and represents the topographic features of a set of points in a landscape.
```{r, echo = FALSE}
showClass("SpatialPointsTopography")
```
+ Class `SpatialGridTopography` extends `SpatialGridDataFrame` and represents the continuous variation of topographic features over a full spatial grid.
```{r, echo = FALSE}
showClass("SpatialGridTopography")
```
+ Class `SpatialPixelsTopography` extends `SpatialPixelsDataFrame` and represents the continuous variation of topographic features over a set if cells in a grid.
```{r, echo = FALSE}
showClass("SpatialPixelsTopography")
```

Although the three classes have the same slots as their parent S4 classes, data frames in `SpatialPointsTopography`, `SpatialGridTopography` and `SpatialPixelsTopography` objects have only three variables: `elevation` (in meters), `slope` (in degrees) and `aspect` (in degrees from North).


### Meteorology
Analogously to topography, three spatial classes are used to represent the variation of daily meteorology over space:

+ Class `SpatialPointsMeteorology` extends `SpatialPoints` and represents daily meteorology series for a set of points in a landscape.
```{r, echo = FALSE}
showClass("SpatialPointsMeteorology")
```
+ Class `SpatialGridMeteorology` extends `SpatialGrid` and represents the continuous variation of daily meteorology across a grid of cells.
```{r, echo = FALSE}
showClass("SpatialGridMeteorology")
```
+ Class `SpatialPixelsMeteorology` extends `SpatialPixels` and represents the variation of daily meteorology for a set of pixels (cells) of a spatial grid.
```{r, echo = FALSE}
showClass("SpatialPixelsMeteorology")
```

In addition to their corresponding inherited slots, `SpatialPointsMeteorology`, `SpatialGridMeteorology` and `SpatialPixelsMeteorology` have two additional slots: `dates` (a vector of days specifying a time period), and `data` (a vector of data frames with the meteorological data). Although the three classes have a `data` slot containing data frames, meteorological data is in different form in each class. In objects of `SpatialPointsMeteorology`, there is one data frame for each point where variables are in columns and dates are in rows. In objects of `SpatialGridMeteorology` and `SpatialPixelsMeteorology`, each data frame describes the meteorology over a complete grid, or a subset of cells, for a single day. In these cases, the data frame has grid cells in rows and variables in columns.


## Reading and writing meteorological data

### Point meteorology
Objects of class `SpatialPointsMeteorology` are stored in the disk using one data file for each of their spatial points. Files can be stored in **ascii** format or **rds** (compressed) format. Package **meteoland** provides four input/output functions for point meteorology:

+ Function `readmeteorologypoint()` reads the meteorological data stored in one *ascii*/*rds* data file and returns a data frame.
+ Function `writemeteorologypoint()` writes the meteorological data of a single point as an *ascii*/*rds* file in the file system.
+ Function `readmeteorologypointfiles()` reads several *ascii*/*rds* files and returns an object of class `SpatialPointsMeteorology`.
+ Functions `writemeteorologypointfiles()` writes several *ascii*/*rds* files in the disk, one per spatial point. Metadata (i.e. the spatial coordinates of each point and the corresponding file path) is stored in an additional file.

### Grid meteorology
 Objects of class `SpatialGridMeteorology` are stored in the disk using one **netCDF** file per day. In this case, the **netCDF** file also contains the date and spatial projection. The following functions are available for input/output of meteorology over a full grid or on a subset of grid cells:

+ Functions `readmeteorologygrid()` and `readmeteorologypixels()` reads the meteorological data stored in one *netCDF* file and returns an object of class `SpatialGridDataframe` or `SpatialPixelsDataframe`, respectively.
+ Functions `writemeteorologygrid()` and `writemeteorologypixels()`  write the meteorological data of the full grid or the subset of grid cells, for a single date in a *netCDF* file.
+ Functions `readmeteorologygridfiles()` and `readmeteorologypixelsfiles()` read several *netCDF* files and returns an object of class `SpatialGridMeteorology` or `SpatialPixelsMeteorology`, respectively.
+ Functions `writemeteorologygridfiles()` and `writemeteorologypixelsfiles()` write several *netCDF* files in the disk, one per date, of the full grid or the subset of grid cells. Metadata (i.e. the dates and their corresponding file path) is stored in an additional file.

As a special case, function `readmeteorologygridcells()` reads several netCDF files and returns an object of class `SpatialPointMeteorology` with the meteorological data of a set of specified grid cells.


## Visualizing input topography and meteorological data

Although very simple, the package provides two kinds of functions to visualize the temporal and spatial variation of meteorology:

+ Function `spplot()` has been redefined from package **sp** to draw maps of specific weather variables corresponding to specific dates. The function can be used on objects of class `SpatialGridMeteorology` and `SpatialPixelsMeteorology`.
+ Function `meteoplot()` allows the temporal series of specific variables on specific spatial points to be plotted. The function can read the data from the disk or from objects of class `SpatialPointsMeteorology`.

Similarly, function `spplot()` also accepts objects of classes `SpatialGridTopography` and `SpatialPixelsTopography`, so that topography can also be easily displayed.


## Manipulating topographic and meteorological data

Following the design of package **sp**, package meteoland provides functions for switching between spatial structures. For example, it is possible to coerce an object of `SpatialPointsTopography` into a `SpatialPixelsTopography` using a call like `as(x, "SpatialPixelsTopography")`. Subsetting is also allowed, so that estimations of daily meteorology can be performed in subsets of points or grids.

The package provides a few functions to extract/reshape meteorological data:

+ Function `extractgridpoints()` extracts the meteorology of particular points in a grid. It accepts objects of `SpatialGridMeteorology` and `SpatialPixelsMeteorology` as input and it returns an object of `SpatialPointsMeteorology`.
+ Function `extractpointdates()` extracts the meteorology of a set of dates from an object of class `SpatialPointsMeteorology`. It returns one data frame for each date, with points in rows and meteorology variables in columns.


## Summarizing meteorological data

The package provides functions to generate temporal summaries of meteorological data. These accept meteorology objects as input and return their corresponding spatial dataframe structure with the summary statistics in columns:

+ Function `summarypoints()` summarizes the meteorology of spatial points. It accepts objects of `SpatialPointsMeteorology` as input and returns an object of `SpatialPointsDataFrame` with point summaries for the requested variable. Temporal summaries can be calculated for different periods and using different summarizing functions (e.g. mean, sum, minimum, maximum, etc.).
+ Functions `summarygrid()` and `summarypixels()` summarize the meteorology of full grids or of subset of grid cells, respectively. They accept objects of `SpatialGridMeteorology` and `SpatialPixelsMeteorology`, respectively, as input and return an object of `SpatialGridDataFrame` and `SpatialPixelsDataFrame`, respectively, with temporal summaries for the requested variable over the range of dates indicated.
+ Function `summaryinterpolationdata()` works similarly to `summarypoints()`, but takes an object of class `MeteorologyInterpolationData` as input.


## Meteorology estimation functions

### Spatial interpolation
Package **meteoland** provides two functions for interpolating meteorological data (i.e., one for each data structure):

+ Function `interpolationpoints()` interpolates weather for a set of locations given in `SpatialPointsTopography` and returns an object of class `SpatialPointsMeteorology`.
+ Function `interpolationgrid()` interpolates weather for a whole grid specified in `SpatialGridTopography` and returns an object of class `SpatialGridMeteorology`.

Both functions require an object of class `MeteorologyInterpolationData`, which contains the X-Y coordinates, the meteorological data and topography of a set of weather stations as well as weather interpolation parameters.
```{r, echo = FALSE}
showClass("MeteorologyInterpolationData")
```
When calling functions `interpolationpoints()` or `interpolationgrid()`, the user may require interpolation outputs to be written into the file system, instead of being returned in memory. If `interpolationpoints()} is called with `export = TRUE}, the function will write the data frame produced for each point into an *ascii* text file or a *rds* file. If `interpolationgrid()` is called with `export = TRUE`, the function will write an *netCDF* file for each day. Metadata files will also be written, so that results can later be loaded in memory. 

Functions `interpolation.calibration()` and `interpolation.cv()` are included in **meteoland** to calibrate interpolation parameters and evaluate predictive performance of interpolation routines before using them.


### Statistical correction

One function is available for statistical correction of meteorological data series (i.e., one function for each data structure). Function `correctionpoints()` performs statistical correction of weather data series on a set of locations and it returns an object of class `SpatialPointsMeteorology` containing corrected weather predictions. Statistical correction requires an object of class `MeteorologyUncorrectedData`, which contains the X-Y coordinates and the coarse-scale meteorological data to be corrected, which includes a reference (historic) period and projected (e.g. future) period:
```{r echo=FALSE}
showClass("MeteorologyUncorrectedData")
```
The reference (historical) period is compared with observed meteorological data of the same period, and the routine uses this information to correct the projected (e.g. future) period. Therefore, apart from the ``MeteorologyUncorrectedData'} object, the correction function requires accurate meteorological data (for a set of spatial points or a grid). Normally, these data will be the result of spatial interpolation.

As before, when calling functions `correctionpoints()`, the user may require the outputs to be written into the file system, instead of being returned in memory. If `correctionpoints()` is called with `export = TRUE`, the function will write the data frame produced for each point into an *ascii* text file or a *rds* file. Metadata files will also be written, so that results can later be loaded in memory.

Function `correctionpoints.errors()` was included in the package to evaluate the errors of the less accurate and more accurate series. Comparisons can be made before and after applying statistical corrections. In the latter case, cross-validation is also available.
